---
title: "Online Retail Analytics"
author: "Franklin Ngochi"
date: "2023-02-26"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plyr)
library(dplyr)
library(tidyr)
library(RcppRoll)
library(scales)
```

#Import the dataset
```{r}
SalesData<-read.csv(file="C:\\Users\\ngoch\\Downloads\\Online_Retail.csv", header=TRUE, sep=",")
colnames(SalesData)
```

#Number and percentage transactions for countries with more than 1% total transactions (Question 1)
```{r}
CountryTransactions<-SalesData%>%count("Country")
names(CountryTransactions)<-c("Country", "Transactions")
CountryTransactions$percentage<-(CountryTransactions[,2]/sum(CountryTransactions[,2]))*100
CountryTransactions
CountryTransactions%>%filter(percentage>"1")
```

#Create Transaction Value variable and add to Dataframe (Question 2)
```{r}
SalesData$TransactionValue<-SalesData$Quantity*SalesData$UnitPrice
head(SalesData, 2)
```

#Show transaction values by country for countries with Total Transaction above 130000 GBP (Question 3)
```{r}
CountryTransactions<-aggregate(SalesData$TransactionValue, list(SalesData$Country), FUN=sum)
names(CountryTransactions)<-c("Country", "TransactionValue")
CountryTransactions%>%filter(TransactionValue>130000)
```

#Question 4 data preparation

```{r}
Temp=strptime(SalesData$InvoiceDate,format='%m/%d/%Y %H:%M',tz='GMT')
head(Temp)
SalesData$New_Invoice_Date<-as.Date(Temp)
SalesData$Invoice_Day_Week=weekdays(SalesData$New_Invoice_Date)
SalesData$Invoice_Day_Hour=as.numeric(format(Temp, "%H"))
SalesData$Invoice_Month=as.numeric(format(Temp, "%m"))
```

#Percentage of transactions by number by days of the week (4a)

```{r}
DailyTransactions<-SalesData%>%count("Invoice_Day_Week")
names(DailyTransactions)<-c("Day", "Transactions")
DailyTransactions$Percentage<-(DailyTransactions[,2]/sum(DailyTransactions[,2]))*100
DailyTransactions
```
#Percentage of transactions by volume by days of the week (4b)

```{r}
DailyTransactionValue<-aggregate(SalesData$TransactionValue, list(SalesData$Invoice_Day_Week), FUN=sum)
names(DailyTransactionValue)<-c("Day", "TransactionVolume")
DailyTransactionValue$Percentage<-(DailyTransactionValue[,2]/sum(DailyTransactionValue[,2]))*100
DailyTransactionValue
```

#Percentage of transactions by volume by month of the year (4c)
```{r}
MonthlyTransactionValue<-aggregate(SalesData$TransactionValue, list(SalesData$Invoice_Month), FUN=sum)
names(MonthlyTransactionValue)<-c("Month", "TransactionVolume")
MonthlyTransactionValue$Percentage<-(MonthlyTransactionValue[,2]/sum(MonthlyTransactionValue[,2]))*100
MonthlyTransactionValue
```

#Date with highest number of transactions for Australia (4d)
```{r}
HighestAustralia<-SalesData%>%filter(Country=="Australia")%>%count("New_Invoice_Date")
names(HighestAustralia)<-c("Date", "Transactions")
HighestAustralia%>%filter(Transactions==max(Transactions))
```
#Best hour to start maintenance (4e)
```{r}
Activity<-SalesData%>%count("Invoice_Day_Hour")#Number of invoices processed each hour
names(Activity)<-c("hour", "InvoicesProcessed")
Activity$PotentialDisturbance<-roll_sum(Activity$InvoicesProcessed,2, fill=NA)#Effect of two-hour interruption on processing per start time
Activity%>%filter(hour%in%(7:20))%>%filter(PotentialDisturbance==min(PotentialDisturbance, na.rm=TRUE))
#start time with the least potential disturbance between 7h00 and 20h00 is 19h00
```

#Plot Histogram for Germany (Question 5)
```{r}
Germany<-select(SalesData, Country,TransactionValue)%>%filter(Country=="Germany")
hist(Germany$TransactionValue, col = 'skyblue3')
```

#Customer with highest number of transactions (Question 6a)
```{r}
SalesData2<-na.omit(SalesData)
TransactionCount<-plyr::count(SalesData2$CustomerID)
names(TransactionCount)<-c("Customer ID", "Transactions")
TransactionCount%>%filter(Transactions==max(Transactions))

```
# Most valuable customer(Question 6b)
```{r}
CustomerValue<-aggregate(SalesData2$TransactionValue, list(SalesData2$CustomerID), FUN=sum)
names(CustomerValue)<-c("CustomerID", "Revenue")
CustomerValue%>%filter(Revenue==max(Revenue))
```

#Percentage of missing values for each variable in dataset (Question 7)
```{r}
colMeans(is.na(SalesData))*100
```

#Number of transactions with missing CustomerID records by countries (Question 8)
```{r}
MissingID<-SalesData%>%filter(is.na(SalesData$CustomerID))
counts<-plyr::count(MissingID$Country)
names(counts)<-c("Country", "Transactions with Missing Customer ID records")
counts
```

#On average, how often the customers return to the website for their next shopping (Question 9)
```{r}


```
#Return rate of the goods purchased by French customers (Question 10)

```{r}
CancelledTransactions<-SalesData%>%filter(Quantity<0)%>%plyr::count("Country")
names(CancelledTransactions)<-c("Country", "CancelledTransactions")
CancelledTransactions$ReturnRate<-CancelledTransactions$CancelledTransactions/nrow(SalesData)
CancelledTransactions%>%filter(Country=="France")


```


#Product that has generated the highest revenue for the retailer(Question 11)

```{r}
ProductValue<-aggregate(SalesData$TransactionValue, list(SalesData$Description), FUN=sum)
names(ProductValue)<-c("Product", "Revenue")
ProductValue%>%filter(Revenue==max(Revenue))
```
#How many unique customers are represented in the dataset (Question 12)
```{r}
paste("The number of Unique customers is", n_distinct(SalesData2[,7]))
```

